## The Pareto Distribution in integer
```{r}
library(MASS)
x<-2:11
alpha<-1:10
beta<-2
f<-beta*alpha^beta/x^(beta+1)
f.new<-paste0("'", fractions(f),"'")

df<-data.frame(cbind(x,alpha,beta,f.new))
print(df)
```


## Illustrate the use of dparetoR function 
```{r}
source("..\\R\\dparetoR.R")
curve( dparetoR( x, 1, 2 ), from = 0, to = 10 )
```


## New table  with f.frac, f.R, and f.Rcpp
```{r}
x<-1:10
alpha=1
beta=2
f<-beta*alpha^beta/x^(beta+1)
f.frac<-paste0("'", fractions(f),"'")

f.R<-dparetoR(x,alpha,beta)

df.new<-data.frame(cbind(x,alpha,beta,f.frac,f.R))
print(df.new)

all.equal(f,f.R)
```


## The code for dparetoR and plots of the density for 3 sets of parameters
```{r}
dparetoR

par(mfrow=c(1,3))
# alpha=1, and beta=2
curve( dparetoR( x, 1, 2 ), from = 0, to = 10 , col='red')
# alpha=2, and beta=3
curve( dparetoR( x, 2, 3 ), from = 0, to = 10 , col='blue')
# alpha=1, and beta=3
curve( dparetoR( x, 1, 3 ), from = 0, to = 10 , col='green')

```


## Compare paretodens function and dpareto function
```{r}
Rcpp::sourceCpp("..\\Rcpp\\dpareto.cpp")

f.Rcpp<-rep(0,10)
for (i in 1:10){
  f.Rcpp[i]=paretodens(i, alpha, beta)
}

f.Rcpp

all.equal(f.R, f.Rcpp)
```

In your written report include the code in an appendix, explain how to load the function and run the program, and show a few (two or three) examples of its use.

```{r, eval = FALSE, code = readLines("dpareto.cpp"), engine = 'Rcpp}
paretodens(2,1,2)
paretodens(3,1,2)
```