---
output:
  pdf_document: default
  html_document: default
---
```{r}
library(MASS)

leuk = MASS::leuk

summary(MASS::leuk )
```


# Optimization
## (a)
The pdf of the life time is a Weibull distribution with density:
$$
f(t_i|\alpha_i, \gamma)=\frac{\gamma}{\alpha_i}(\frac{t_i}{\alpha_i})^{\gamma-1}exp(-(\frac{t_i}{\alpha_i})^{\gamma})
$$
with $log(\alpha_i)=\beta_o+\beta_1x_i+\beta_2u_i$.
Let's assume $\delta= 1/\gamma$ and $z_i=(logt_i-\beta_o-\beta_1x_i-\beta_2u_i)/\delta$, then:
$$
\gamma = 1/\delta\\
 \gamma log(\frac{t_i}{\alpha_i})=z_i
$$
The log likelihood of for this model is:

$$
l(\alpha_i, \gamma)=logf=log\gamma-log\alpha_i+\gamma log(\frac{t_i}{\alpha_i})-log(\frac{t_i}{\alpha_i})-(\frac{t_i}{\alpha_i})^{\gamma}\\
=-log\delta-log\alpha_i+z_i-logt_i+log\alpha_i-exp(z_i)\\
=-log\delta-logt_i+z_i-exp(z_i)\\
l(z, \delta)=-nlog\delta-\sum_ilogt_i+\sum_i(z_i-exp(z_i))
$$
After dropping additive terms constant in the parameters,this log-likelihood can be written as:
$$
l(\beta_0, \beta_1,\beta_2,\delta)=-nlog\delta+\sum_i(z_i-exp(z_i))
$$

## (b)
```{r}
t <- leuk$time
x <- log(leuk$wbc/10000)
u <- ifelse(leuk$ag == "present", 1, 0)


# Define log-likelihood function
loglik <- function(para) {
  beta0 <- para[1]
  beta1 <- para[2]
  beta2 <- para[3]
  delta <- para[4]
  z <- (log(t) - beta0 - beta1 * x - beta2 * u) / delta
  res = sum(z - exp(z)) - length(t) * log(delta) - sum(log(t))
  return(-res)
}
```

```{r, warning=FALSE}
# Find MLEs using optim
start_para <- c(1, 3, 1, 2)
optim_out <- optim(start_para, loglik, hessian = TRUE)
beta0_mle <- optim_out$par[1]
beta1_mle <- optim_out$par[2]
beta2_mle <- optim_out$par[3]
delta_mle <- optim_out$par[4]

vcov_mle <- solve(optim_out$hessian)

# Compute standard errors
se_beta0_mle <- sqrt(vcov_mle[1, 1])
se_beta1_mle <- sqrt(vcov_mle[2, 2])
se_beta2_mle <- sqrt(vcov_mle[3, 3])
se_delta_mle <- sqrt(vcov_mle[4, 4])


mle_table <- data.frame(Parameter = c("beta0", "beta1", "beta2", "delta"),
                         MLE = c(beta0_mle, beta1_mle, beta2_mle, delta_mle),
                         SE = c(se_beta0_mle, se_beta1_mle, se_beta2_mle, se_delta_mle))
print(mle_table)
```


## (c)
Take partial derivative to each parameters:
$$
\frac{\partial l}{\partial\beta_0}=-\frac{n}{\delta}+\frac{1}{\delta}\sum_iexp(\frac{logt_i-\beta_0-\beta_1x_i-\beta_2u_i}{\delta})\\
$$
$$
\frac{\partial l}{\partial\beta_1}=-\frac{\sum_ix_i}{\delta}+\frac{1}{\delta}\sum_ix_iexp(\frac{logt_i-\beta_0-\beta_1x_i-\beta_2u_i}{\delta})\\
$$
$$
\frac{\partial l}{\partial\beta_2}=-\frac{\sum_iu_i}{\delta}+\frac{1}{\delta}\sum_iu_iexp(\frac{logt_i-\beta_0-\beta_1x_i-\beta_2u_i}{\delta})\\
$$
$$
\frac{\partial l}{\partial\delta}=-\frac{n}{\delta}-\frac{\sum_i(logt_i-\beta_0-\beta_1x_i-\beta_2u_i)}{\delta^2}+\frac{1}{\delta^2}\sum_i(logt_i-\beta_0-\beta_1x_i-\beta_2u_i)exp(\frac{logt_i-\beta_0-\beta_1x_i-\beta_2u_i}{\delta})
$$
Then replace each with the estimated value we have from the optim function:
```{r}
n = length(t)
z = (log(t) - beta0_mle - beta1_mle * x - beta2_mle * u) / delta_mle
beta0_partial = - n/delta_mle + 1/delta_mle * sum(exp(z))
beta1_partial = -sum(x)/delta_mle + 1/delta_mle * sum(x*exp(z))
beta2_partial = -sum(u)/delta_mle + 1/delta_mle * sum(u*exp(z))
delta_partial = - n/delta_mle - sum(z)/(delta_mle) + 1/(delta_mle^2) * sum(z*delta_mle*exp(z))


beta0_partial
beta1_partial
beta2_partial
delta_partial
```
Partial derivatives of each parameter are close to zero. 

```{r}
library(numDeriv)
para_mle = c(beta0_mle, beta1_mle, beta2_mle, delta_mle)
grad(loglik,para_mle, method = "complex")
```
Partial derivatives of each parameter are close to zero. 



